---
title: "Homework 2"
author: "Sarahy Martinez"
date: "09-29-2024"
output: github_document
---

Libraries Used for this HW
```{r message=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
```

## Problem 1


```{r message=FALSE}
 NYC_transit =
  read_csv("./data/NYC_Transit_Subway_Entrance_AND_EXIT_DATA.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  relocate(line:entry, ada,vending) %>% 
  select(-division:-entrance_location, station_location) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE)) %>% 
  mutate(vending = ifelse(vending == "YES", TRUE, FALSE)) %>% 
  mutate(across(starts_with("route"), as.character)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name",
    values_drop_na = TRUE
  ) %>% 
  relocate(station_name,station_location,route_number, route_name,line,entrance_type, entry, ada,vending)

```

This dataset contains the variables  station name, station location, train route number, routes,line, entrance type, entry, exit only, vending, ada, station lat, and station long. I cleaned my data by using clean names to fix the variable names, relocated the order of my variables for legibility, selected my columns/ data of interest, mutated the character variables from entry to logical, mutates the names of the routes as characters because it wasn't allowing me to run the pivot function , then used the pivot_longer to tidy the data and combine route numbers to one column  and assign the routes their own column with the station, and relocated again for legibility.  Data is now tidy.



```{r}
count(distinct(NYC_transit, station_name,line))

count(filter(NYC_transit, ada == TRUE))


count(filter(NYC_transit, entry == TRUE, vending == FALSE))

A_train =
  NYC_transit %>% 
  filter(route_name == "A") %>% 
  distinct(station_name) 
nrow(A_train)

A_train_ada =
  NYC_transit %>% 
  filter(route_name == "A", ada== TRUE) %>% 
  distinct(station_name) 
nrow(A_train_ada)
 
```

* There are 465 distinct stations.
* There are 1,616 stations that are ADA compliant. 
* There are a total of 139 stations entrances/ exits without vending that allow entrance, its proportion is 1. 
* There are 56 stations that serve the A train, and 16 are ADA compliant. 


## Probem 2


Impporting and reading the dataset
```{r}

mister_wheel_data = read_excel("./data/202309 Trash Wheel Collection Data.xlsx")

```



























