---
title: "Homework 2"
author: "Sarahy Martinez"
date: "09-29-2024"
output: github_document
---

Libraries Used for this HW
```{r message=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
```

## Problem 1


```{r, cleaning_nyc, message=FALSE}
 NYC_transit =
  read_csv("./data/NYC_Transit_Subway_Entrance_AND_EXIT_DATA.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  relocate(line:entry, ada,vending) %>% 
  select(-division:-entrance_location, station_location) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE)) %>% 
  mutate(vending = ifelse(vending == "YES", TRUE, FALSE)) %>% 
  mutate(across(starts_with("route"), as.character)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name",
  ) %>% 
  relocate(station_name,station_location,route_number, route_name,line,entrance_type, entry, ada,vending)

```

This dataset contains the variables  station name, station location, train route number, routes,line, entrance type, entry, exit only, vending, ada, station lat, and station long. I cleaned my data by using clean names to fix the variable names, relocated the order of my variables for legibility, selected my columns/ data of interest, mutated the character variables from entry to logical, mutates the names of the routes as characters because it wasn't allowing me to run the pivot function , then used the pivot_longer to tidy the data and combine route numbers to one column  and assign the routes their own column with the station, and relocated again for legibility.  Data is now tidy.



```{r, counting}
count(distinct(NYC_transit, station_name,line))

transit_entry = 
  NYC_transit %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct()

transit_entry = NYC_transit %>% 
  filter( vending == "FALSE") %>% 
  pull(entry) %>% 
  mean()

A_train =
  NYC_transit %>% 
  filter(route_name == "A") %>% 
  distinct(station_name, line) 
nrow(A_train)

A_train_ada =
  NYC_transit %>% 
  filter(route_name == "A", ada== TRUE) %>% 
  distinct(station_name, line) 
nrow(A_train_ada)
 
```

* There are 465 distinct stations.
* There are 84 stations that are ADA compliant. 
* There proportion of entrances/ exits without vending that allow entrance, is  0.377. 
* There are 60 stations that serve the A train, and 17 are ADA compliant. 


## Probem 2


Importing and reading the dataset

```{r, mister_trash}

mister_trash_data = read_excel("./data/202409 Trash Wheel Collection Data New.xlsx", 
                               sheet=1,
                               skip = 1,
                               range= "A2:N653",
                               na = c("NA", ".", "")) %>% 
                            janitor::clean_names() %>% 
                            mutate(sports_balls = as.integer(round(sports_balls))) %>% 
                            mutate(year = as.character(year)) %>% 
                            mutate(trash = "mister_trash") %>% 
                            drop_na()


```




```{r, professor_trash}

professor_trash = read_excel("./data/202409 Trash Wheel Collection Data New.xlsx", 
                               sheet=2,
                               skip = 1,
                               range= "A2:M121",
                               na = c("NA", ".", "")) %>% 
                            janitor::clean_names() %>% 
                            mutate(year = as.character(year)) %>% 
                            mutate(trash = "professor_trash") %>% 
                            drop_na()


```



```{r, gwynnda_sheet}

gwynnda_trash = read_excel("./data/202409 Trash Wheel Collection Data New.xlsx", 
                               sheet=4,
                               skip = 1,
                               range= "A2:L265",
                               na = c("NA", ".", "")) %>% 
                            janitor::clean_names() %>% 
                            mutate(year = as.character(year)) %>% 
                            mutate(trash = "gwynnda_trash") 
                            

```


```{r, combined, message=FALSE}

trash_wheel_tidy = 
  bind_rows(professor_trash, gwynnda_trash, mister_trash_data) %>% 
  janitor::clean_names() %>% 
  select(trash, everything())


```

In this dataset we have 15 variables which include dumpster, month, year, date, weight_tons, volume_cubic_yards, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, plastic_bags, wrappers, homes_powered,sports_balls, and trash (to specify from which dataset). There are a total of `r nrow(trash_wheel_tidy)` observations and `r ncol(trash_wheel_tidy)` columns. The datasheets were individually imported, in mister trash I mutated the year and same for the rest due to years being considered as a character in one dataset and the others as numeric so I converted them all to characters. I also cleaned the names for reasonable variable names and mutated again to create a variable name to identify which dataset the trash is from. 

```{r, caclulate,message=FALSE}

total_weight_prof =  trash_wheel_tidy %>% 
                     filter(trash == "professor_trash") %>% 
                     summarise(total_weight = sum(weight_tons, na.rm = TRUE))



total_cigs_gwn =  trash_wheel_tidy %>% 
                     filter(trash == "gwynnda_trash", year == "2022", month == "June") %>% 
                     summarise(cigarette_butts = sum(cigarette_butts))

```

* The total weight of trash collected by Professor trash wheel was `r (total_weight_prof)`. 
* In the month of June 2022 Gwynnda collected `r options(scipen = 999);total_cigs_gwn` cigarette butts. 


## Problem 3

```{r, original_bakes}


bakers_df = 
  read_csv("./data/bakers.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  separate(baker_name, into = c("first_name", "last_name"), sep = " ") %>% 
  select(- last_name) %>% 
  rename(bakers = first_name) %>% 
  mutate( bakers = ifelse(bakers == "Jo","Joanne", bakers)) %>% 
  drop_na()

bakes_df = 
  read_csv("./data/bakes.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  rename(bakers = baker) %>% 
    mutate(bakers = ifelse(bakers == '"Jo"', "Joanne", bakers)) %>% 
  drop_na()

results_df = 
  read_csv("./data/results.csv", skip = 3, col_names = FALSE,na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  rename(series = x1,episode = x2, baker = x3, technical = x4, result = x5, ) %>% 
  rename( bakers = baker) %>% 
  drop_na()

anti_join(results_df, bakes_df, by = c("bakers", "series", "episode"))

 anti_join(results_df, bakers_df, by = c("series","bakers"))
 
  GBB_merged_ = 
  full_join(results_df, bakes_df, by = c("bakers", "series", "episode")) %>% 
  full_join(x= ., bakers_df, by =c("series","bakers")) %>% 
  select(series, episode, bakers, baker_age, technical,result, baker_occupation, hometown, signature_bake, show_stopper)

write_csv(GBB_merged_,"./data/GBB_new.csv")

```
For each of the datasets I imported them through read_csv and then cleaned the names to follow naming conventions. In the bakers files I split the first and last name because they do not follow normal form and removed the last name since the other datasets do not include them.I then renamed baker to bakers just for legible purposes and noticed that there was a Jo and Joanne, while searching online noticed that is her nickname so I renamed it to Joanne by mutating. Same process follows for the bakes and results except in the results I had to rename my variables by mutating them. I used an antijoin  to note any discepancies and noticed it was Jo, so I changes the names to Joanne, in the other anti-join there seems to be descrepancies but I am assuming its becuase of the occupations not being involved in the results. I then merged my datasets by doing a full join and selected my variables so that they can be arranged in a way that makes sense when reading the CSV. 

The final dataset written as GBB_merged_ and it contain the variables series, episode, bakers, baker_age, technical, result, baker_occupation, hometown, signature_bake, show_stopper. The total observations `r nrow(GBB_merged_)` and columns `r ncol(GBB_merged_)`

```{r, combined, stars_merge}

 GBB_merged_ = 
  full_join(results_df, bakes_df, by = c("bakers", "series", "episode")) %>% 
  full_join(x= ., bakers_df, by =c("series","bakers")) %>% 
  select(series, episode, bakers, baker_age, technical,result, baker_occupation, hometown, signature_bake, show_stopper)


GBB_merged_star = 
 GBB_merged_  %>% 
  filter(series >= 5 & series <= 10) %>% 
  select(series, episode,bakers, result) %>% 
  filter(result %in% c("WINNER", "STAR BAKER")) %>% 
  distinct( )  %>% 
  knitr:: kable()

```

```{r, combined, stars}

print(GBB_merged_star)

```


Candice and Nadiya were predictable overall winners because they were STAR BAKERS more than once. David was a surprise for winning in season 10 since he never was crowned STAR BAKER. 

## Problem 3, part 2 

```{r, views, message=FALSE}
views = 
  read_csv("./data/viewers.csv",na = c("NA", ".", ""))  %>% 
  janitor::clean_names() %>% 
  rename( season_1 = series_1, season_2 = series_2, season_3 = series_3, season_4 = series_4, season_5 = series_5, season_6 = series_6, season_7 = series_7, season_8 = series_8, season_9 = series_9, season_10 = series_10) %>% 
  head(10)
 


```

* The average viewership in season 1 was `r mean(pull(views, season_1), na.rm = TRUE) ` .
* The average viewership in season 5 was `r mean(pull(views, season_5), na.rm = TRUE)`. 










