---
title: "Homework 2"
author: "Sarahy Martinez"
date: "09-29-2024"
output: github_document
---

Libraries Used for this HW
```{r message=FALSE}
library(tidyverse)
library(tidyr)
library(readxl)
```

## Problem 1


```{r, cleaning_nyc, message=FALSE}
 NYC_transit =
  read_csv("./data/NYC_Transit_Subway_Entrance_AND_EXIT_DATA.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  relocate(line:entry, ada,vending) %>% 
  select(-division:-entrance_location, station_location) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE)) %>% 
  mutate(vending = ifelse(vending == "YES", TRUE, FALSE)) %>% 
  mutate(across(starts_with("route"), as.character)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name",
    values_drop_na = TRUE
  ) %>% 
  relocate(station_name,station_location,route_number, route_name,line,entrance_type, entry, ada,vending)

```

This dataset contains the variables  station name, station location, train route number, routes,line, entrance type, entry, exit only, vending, ada, station lat, and station long. I cleaned my data by using clean names to fix the variable names, relocated the order of my variables for legibility, selected my columns/ data of interest, mutated the character variables from entry to logical, mutates the names of the routes as characters because it wasn't allowing me to run the pivot function , then used the pivot_longer to tidy the data and combine route numbers to one column  and assign the routes their own column with the station, and relocated again for legibility.  Data is now tidy.



```{r, counting}
count(distinct(NYC_transit, station_name,line))

count(filter(NYC_transit, ada == TRUE))


count(filter(NYC_transit, entry == TRUE, vending == FALSE))

A_train =
  NYC_transit %>% 
  filter(route_name == "A") %>% 
  distinct(station_name) 
nrow(A_train)

A_train_ada =
  NYC_transit %>% 
  filter(route_name == "A", ada== TRUE) %>% 
  distinct(station_name) 
nrow(A_train_ada)
 
```

* There are 465 distinct stations.
* There are 1,616 stations that are ADA compliant. 
* There are a total of 139 stations entrances/ exits without vending that allow entrance, its proportion is 1. 
* There are 56 stations that serve the A train, and 16 are ADA compliant. 


## Probem 2


Importing and reading the dataset

```{r, mister_trash}

mister_wheel_data = read_excel("./data/202309 Trash Wheel Collection Data.xlsx", 
                               sheet=1,
                               skip = 1,
                               range= "A2:N586",
                               na = c("NA", ".", "")) %>% 
                            janitor::clean_names() %>% 
                            mutate(sports_balls = as.integer(round(sports_balls))) %>% 
                            mutate(year = as.numeric(year)) %>% 
                            mutate(trash = "mister_trash") %>% 
                            na.omit()
```




```{r, professor_trash}

professor_trash = read_excel("./data/202309 Trash Wheel Collection Data.xlsx", 
                               sheet=2,
                               skip = 1,
                               range= "A2:M108",
                               na = c("NA", ".", "")) %>% 
                            janitor::clean_names() %>% 
                            mutate(year = as.numeric(year)) %>% 
                            mutate(trash = "professor_trash") %>% 
                            na.omit()


```



```{r, gwynnda_sheet}

gwynnda_trash = read_excel("./data/202309 Trash Wheel Collection Data.xlsx", 
                               sheet=4,
                               skip = 1,
                               range= "A2:L157",
                               na = c("NA", ".", "")) %>% 
                            janitor::clean_names() %>% 
                            mutate(year = as.numeric(year)) %>% 
                            mutate(trash = "gwynnda_trash") %>% 
                            na.omit()


```


```{r}

trash_wheel_tidy = 
  bind_rows(professor_trash, gwynnda_trash, mister_wheel_data) %>% 
  janitor::clean_names() 

nrow(trash_wheel_tidy)
ncol(trash_wheel_tidy)
summary(trash_wheel_tidy)

# with na.omit 

trash_wheel_tidy_omit =
   bind_rows(professor_trash, gwynnda_trash, mister_wheel_data) %>% 
  janitor::clean_names() %>% 
  na.omit
  
nrow(trash_wheel_tidy_omit)
ncol(trash_wheel_tidy_omit)
summary(trash_wheel_tidy_omit)
```
In this dataset we have the variables dumpster, month, year, date, weight_tons, volume_cubic_yards, plastic_bottles, polystyrene, cigarette_butts, glass_bottles, plastic_bags, wrappers, homes_powered,sports_balls, and trash (to specify from which dataset). There are a total of 845 observations (rows) and 15 columns without omitted data and a total of 584 rows (observations) and  15 columns with omitted data. The datasheets were individually imported, in mister trash I mutated the year and same for the rest due to years being considered as a character in one dataset and the others as numeric so I converted them all to numeric and get the mean years, committed data is 2018 and non- omitted is 2019. I also cleaned the names for reasonable variable names and mutated again to create a variable name to identify which dataset the trash is from. 

```{r}

total_weight_prof_omit =  trash_wheel_tidy_omit %>% 
                     filter(trash == "professor_trash") %>% 
                     summarise(total_weight = sum(weight_tons, na.rm = TRUE))

print(total_weight_prof_omit)
#with ommitted data total weight is 0 because professor_trash is not in this dataset with omitted data for removing info that does not have trash specific data.


total_weight_prof =  trash_wheel_tidy %>% 
                     filter(trash == "professor_trash") %>% 
                     summarise(total_weight = sum(weight_tons, na.rm = TRUE))

print(total_weight_prof)


total_cigs_gwn =  trash_wheel_tidy %>% 
                     filter(trash == "gwynnda_trash", year == 2022, month == "June") %>% 
                     summarise(cigarette_butts = sum(cigarette_butts, na.rm = TRUE))
print(total_cigs_gwn)

```


* The total weight of trash collected by Professor trash wheel was 210.11 tons (without entire binded dataset omitted). With the omitted dataset professor trash collected 0 tons.
* In the month of June 2022 Gwynnda collected 0 cigarette butts. 


## Problem 3


```{r}
bakers = 
  read_csv("./data/bakers.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  separate(baker_name, into = c("first_name", "last_name"), sep = " ") %>% 
  select(- last_name) %>% 
  rename(baker = first_name)
  
```



```{r, bakes}
bakes = 
  read_csv("./data/bakes.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names()

print(bakes)
```


```{r, results}
results = 
  read_csv("./data/results.csv", skip = 3, col_names = FALSE,na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  rename(series = x1,episode = x2, baker = x3, technical = x4, result = x5 )

print(results)

```



```{r, views}
views = 
  read_csv("./data/viewers.csv",na = c("NA", ".", "")) %>% 
  janitor::clean_names() %>% 
  head(10) %>% 
   mutate(across(c(series_1:series_10), as.numeric))



```












